Impacts of the AUMF Resolutions on Obama and Trump Administration Foreign Policy

### 2020-12-07 paper.tex
first draft complete

It seems as if the changes in foreign policy have been more along the lines of application of prior legal interpretations rather than substantial changes in the interpretations themselves.
President Trump has also continued the use of drone strikes outside of the original 9/11 theaters, as was seen in the Obama presidency \autocite[35]{sterio2018}.
Along these lines, the Trump Administration has legally agreed with the prior Bush and Obama administration interpretations of the 2001 AUMF in regards to drone strikes.
The willingness to use lethal force on suspected terrorists, as shown by both the drone strikes as well as specific incidents such as the assassination of Iran's General Qassem Soleimani, demonstrate that Trump has been more than willing to stick to the Bush Administrations armed conflict paradigm instead of attempting to shift approaches like the Obama Administration.
The Trump administration the covert and secretive nature of force application that was found in both of the prior administrations.

>>>>

It seems as if the changes in foreign policy have been more along the lines of application of prior legal interpretations rather than substantial changes in the interpretations themselves.
President Trump has also continued the use of drone strikes outside of the original 9/11 theaters, as was seen in the Obama presidency \autocite[35]{sterio2018}.
Along these lines, the Trump Administration has legally agreed with the prior Bush and Obama administration interpretations of the 2001 AUMF in regards to drone strikes.
The willingness to use lethal force on suspected terrorists, as shown by both the drone strikes as well as specific incidents such as the assassination of Iran's General Qassem Soleimani, demonstrate that Trump has been more than willing to stick to the Bush Administrations armed conflict paradigm instead of attempting to shift approaches like the Obama Administration.
The Trump administration also maintained the covert and secretive nature of executive military force that was found in both of the prior administrations.

\section*{Conclusion}
Throughout the course of this research project, the main thing I learned is how complex the history and discussion around the 2001 AUMF is.
I had no idea how much effort and detail went into the drafting of the resolution, and also how widely it has been applied to the foreign policy of the United States for the last two decades.
I also believe that I have learned a lot about the foreign policy stances of the Bush, Obama, and Trump Administrations in general.
Although I had heard a lot of news and talk about their foreign policy decisions through current events, speeches, and talking with others, it was interesting to see the ways in which the public image of the presidents seemed to in many cases conflict with the actually legal and political actions taken by their administrations.
For example, the common rhetoric around the Trump presidency might suggest that he has radically shifted our military policy away from that of the Obama Administration (regardless of whether that is a good or bad thing).
Although this was true to an extent, the changes largely seemed to be in his rhetoric and application rather than any kind of legal shift.
This is similar to the talk around the Obama Administration, as it seems that many people don't have an accurate understanding of the actions President Obama regarding executive military power.
While many of his liberal defenders seem to believe that he was able to undo the wrongs of the Bush Administration, many of his opponents on both the left and the right have criticized him for preserving the status quo.
In reality, it seems to be a mix between those stances.
I believe that while President Obama stayed similar in his policy and legal decisions to the President Bush, there were some notable changes in legal frameworks that could lead to positive change in the future.
However, I also believe that it is impossible to overlook the negative sides of Obama's AUMF application whether it be the secretive drone program, the preservation of Guantanamo Bay, or the global expansion of the war on terror.

In regards to what I believe should happen, I am now a firm believer in the stance that the three previous AUMF resolutions ought to be repealed.
This is because of two main reasons.
First, their existence has stretched executive overreach far beyond what was imagined by the original authors of the 2001 AUMF resolution.
It is clear from the actions taken by Congress in the drafting process that they wished for a one-time authorization rather than a more general blank check for executive branch military force.
It is also clear that the scope of the authorization was intended to just those who were directly involved in the specific attacks of 9/11, not terrorism as a whole.
I believe that expansions of the AUMFs to include groups such as Al-Shabaab and ISIS is a blatant abuse of power on the side of the Bush, Obama, and Trump administrations.
Second, I believe that their use has negatively affected transparency and accountability within our government.
We saw with the prior presidential administrations that the AUMF was used as a justification for countless instances of covert action, which seems to completely contradict the original meaning of the resolution.
Although the resolutions try to provide oversight through the invocation of the War Power Resolution, it is clear that this oversight has not been achieved as neither Congress nor the American public were informed about the true extent of the war on terror.

Although oversight and limitations could theoretically be implemented through different legal interpretations of the 2001 AUMF rather than completely new legislation, it seems to me that the approach of repealing the resolution would still be preferable.
Due to the \emph{fait accompli} nature of the war on terror that was brought up by Starr-Deelen, I don't see the possibility of future presidents being able to significantly change the legal interpretations of that arose in the Bush and Obama administrations.
As there are currently conflicts that are being passed down from president to president, it is getting to the point where it is much easier for the president to manage the wars rather than completely end them.
Therefore, any kind of change in the future will most likely have to come from either the legislative or judicial branches of our government.
Since the judicial branch is full of appointees from the prior administrations, and because they seemed to largely agree with the interpretations of the Obama administration, there isn't much hope for change in that branch.
As a result, repealing the 2001 AUMF not only seems to be the most preferable approach but also the most realistic approach.

In terms of what is actually being done within Congress right now, there seems to be some kind of willingness to repeal the AUMF resolution, even if it isn't a top priority.
Although there have been many attempts to pass resolutions to repeal the authorizations, they have largely been unsuccessful.
Tim Kaine, among with other Democratic members of Congress, have led the fight against the AUMF with the main opposition being from Republicans.
Senator McConnell has repeatedly blocked resolutions in Senate that would address the AUMFs, and has also spoken about how the United States is not doing enough to combat the rise of ISIS\@.
The closest that Congress has come to a repeal was in 2019, when the House Democrats included a small section in an appropriations bill that would repeal the 2001 AUMF resolution.
Although this did not end up passing through the Senate, it was the first time that any legislation has been passed to repeal an AUMF resolution \autocite{fuller2019}.
There has also been slight support from certain Republicans.
Senator Rand Paul of Kentucky has stated that ``it is indisputably clear that the authority given to the executive under the AUMF passed after 9/11 has become too broad and needs updating \autocite{paul2018}.''
Although it has been almost two decades since the 2001 AUMF was passed, the conversation is far from over.
The time to properly address its problems seems to be getting closer every day.
